### 2012-12-03 write_up.tex
Added write up template / added latex files to .gitignore

\documentclass{article}

\title{Paper}
\author{T. Martian \& E. Brinkbot}

\begin{document}

\maketitle

\section{Complexity}

working on this section...

\end{document}