### 2012-12-03 write_up.tex
Did work on write up. Added Model section, Algorithm section, Complexity section

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{Paper}
\author{T. Martian \& E. Brinkbot}

\begin{document}

\maketitle

\section{Model}

Model Description goes here...

\begin{align*}
  & k && \text{Number of blocks} \\
  & \{s_i\}_{i=1..k} && \text{Number of nodes in each block}
\end{align*}

\section{Algorithm}

The algorithm uses two lookup tables (one for the scheduler, and one
for the nodes) to determine the optimal play given the relevant
statistics of the game at decision time. Each table has a dimension
for every relevant play statistic for the deciding agent, and contains
their optimal choice as well as the expected number of Yeses in each
block after that choice is made (which is necessary for callers to
decide on their optimal decision, as expected payoffs are only a
function of the expected number of Yeses at the end).

These two tables back two oracles (wrong terminology?) which will
return a precomputed value if it's already in the table, or they'll
compute it as defined below. This can lead to a recursive cascade of
computation, which is bounded by the size of the tables times the time
to compute each element in terms of a lookup in one of them.

We'll call the scheduler oracle
\begin{align*}
  & \operatorname{sched}(\{c_i\}, \{y_i\})
  \intertext{and the node oracle}
  & \operatorname{node}(t, b, \{c_i\}, \{y_i\})
  \intertext{where}
  & type && \text{Node's type (Yes or No)} \\
  & n && \text{Node block} \\
  & \{c_i\} && \text{Number of already chosen nodes by block} \\
  & \{y_i\} && \text{Number of Yes chosen nodes by block}
\end{align*}
We can show (we need to show) that that's all of the relevant
statistics either agent needs to know to determine a unique optimal
decision.

The scheduler just needs to check each possible choice of next block,
and for each block calculate the expected number of Yeses by weighting
the expected number of Yeses for a Yes or No node by the probability
of a Yes or No node. This involves at most $2k$ calls to the node
oracle, and is sufficient for determining the optimal choice of the
scheduler.

The nodes just have to find their expected value if they choose Yes or
choose No. This is just comparing two calls to the sched oracle.

Combining these two yields a polynomial in $n$ time algorithm for
solving arbitrary block models.

\section{Complexity}

lookups are constant time, so the total time complexity is going to
be bounded by the number of entries in the table times the number of
table lookups it takes to compute one entry in the table. Later we may
be able to show not all of the table needs / will be computed lowering
complexity.

The size of the node table is trivial to upper bound. There are $2k+2$
dimensions: the node's type (2 values), the node's block ($k$ values),
$k$ dimensions for the number of nodes in each block that have gone
before ($\le s_i + 1 | i=1..k$), and $k$ dimensions for the number of
nodes in each block that have gone before and selected yes ($\le s_i +
1 | i=1..k$). This leads to the following statement,
\begin{align*}
  NodeTableSize &= o\left( 2k \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( k \prod_{i=1}^k s_i^2 \right)
  \intertext{If we further assume that the size of $\ell \le k$ of
    the blocks are order $n$ and the rest are constant size, then the
    size of the table reduces to}
  NodeTableSize &= o\left( kn^{2\ell} \right) \footnotemark
  \intertext{Next we need to find the complexity of the computing one
    entry in the node table in terms of other table lookups. Each
    entry computation takes 2 calls to the sched oracle, and then does
    computation of the expected number of Yeses which is order $k$,
    making the total complexity for the node table given the sched
    table}
  NodeTableSize &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  NodeTableSize &= o\left( k^2 n^{2\ell} \right)
  \intertext{The size of the scheduler table is also trivial to upper
    bound. There are $2k$ dimensions similar to the node table. By
    similar reasoning}
  SchedTableSize &= o\left( \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( \prod_{i=1}^k s_i^2 \right)
  \intertext{If we make the same assumption about block sizes this
    reduces to}
  SchedTableSize &= o \left( n^{2\ell} \right)
  \intertext{The complexity for each entry in the scheduler table is
    linear in $k$, because it makes $2k$ lookups to the node
    table. Therefore the total complexity for constructing the
    scheduler table is bounded by}
  SchedTableSize &= o\left( k \prod_{i=1}^k s_i^2 \right) \\
  SchedTableSize &= o \left( k n^{2\ell} \right)
  \intertext{This leaves us with final time complexity of}
  RunningTime &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
\end{align*}
\footnotetext{Note, this may be smaller $o\left( \ell n^{2\ell}
  \right)$, not sure}

Similarly the node table is the large of the two tables, and has space
complexity equal to time complexity.

\end{document}