### 2012-12-07 proposal.tex
Added files for proposal to help with paper

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Proposal: Studying strategic cascades using 2-blockmodels}
\author{T. Martin \& E. Brinkman}
\date{\today}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{fullpage} %my stuff was already two pages at standard margins

\begin{document}

\maketitle

\section{Introduction}
\label{intro}

The study of network games is a relatively new and very active area of research. It has become both more important and more applicable since the internet has made large scale social interaction easier and more quantifiable. A common topic in network games is that of contagious or cascading processes, in which a number of agents start with some property they then spread to their neighbors under some spreading rules. This model naturally represents phenomena such as the spreading of trends, technologies, or influence through people or groups, or cascading failures in structures such as power grids or banks. Our proposed research is about theoretical models of spreading and cascading processes on networks.

Many models with simple spreading rules have been proposed \cite{Arthur89, Morris00, Watts02} to explain, for example, how breaking news spreads over the internet or how a new technology (such as the iPod) spreads in popularity. These spreading rules utually take one of two forms. The rules either govern a stochastic spreading process, or they assume that each node in the network is a strategic agent which makes a decision (e.g. to buy an iPod or to buy a Zune) and derives greater utility if his neighbors make the same decision. To the best of our knowledge, all current strategic agent models make a simplifying assumption that the agents are only myopically strategic \cite{Chierichetti12}: they make decisions at each round to maximize their utility at that point in the game. But, each agent's behavior can have profound impact on the future state of the network and thus the agent's own final utility. Our interest is in studying network processes under the more realistic assumption that agents make decisions conscious of their future influence on outcomes, in order to maximize their expected final utility.

More specifically, we propose to research the impact of having strategic users: how much different are results if we assume users are strategic instead of myopic. To answer this question, we will use a model introduced in \cite{Chierichetti12} and described in Section~\ref{prob_statement}. Our contribution is to fully characterize the behavior of this model on a specific class of graphs, the 2-blockmodels, in hopes of learning more about the general case. Blockmodels have been studied extensively in the past \cite{Wang87, Snijders97} as a natural framing of networks in which we can think of a node having a type which influences its characteristics. For example, a blockmodel describing a social network could have types for each grade, and students with each type are more likely to connect to their own grade, and unlikely to connect to grades very far away.

\section{Problem Statement}
\label{prob_statement}
We propose studying a simple cascade model over an arbitrary social network. In this model each agent is represented by a node in a graph and all of their friends / neighbors are represented by edges in the same graph. Over the course of the game each agent will make an irreversible choice between one of two types ($Y$ or $N$) -- where agents that prefer $Y$ are less frequent -- and will get a payoff at the end of the game corresponding to the number of neighbors that chose the same type as well as a bonus if they chose their prefered type. Thus each agent is torn between choosing their prefered type and choosing the type that a majority of their neighbors chose. The game is played with a single node making a choice between types at each point in time. We are looking to choose an order or schedule over agents to maximize the number of minority ($Y$) choices. We are interested in two different classes of schedules, offline and online. Offline schedules must schedule the entire set of agents before the game is played, whereas online schedules can be revised as agents make their decisions. We are interested in the difference between the optimal number of minority types chosen when the agents behave myopically, meaning their choice optimizes their current payoff, and when the agents behave strategically, meaning their choice optimizes their expected payoff at the end of the game. Strategic agents vary significantly depending on how much information they have. Here we assume they know the entire graph structure, every other agents strategy, the probability of an agent’s type, all decided node's choices, and the schedule.

Formally the model is specified for a graph $G(V,E)$ where each agent is a node $v_i \in V, i={1..n}$. At every point in time $t={0..n}$ each agent is in one of three states, Yes, No, or Undecided ($v_{it} \in \{Y, N, U\}$). At time zero each agent starts off Undecided, then at each point in time one agent according to the schedule switches from Undecided to either Yes or No. The schedule is an ordered permutation ($\mathcal P$) of the nodes. For the offline schedule this is static a permutation chosen ahead of time, but in the online schedule the $j^{th}$ element of the permutation is a function of the graph after the previous time. If the permutation is definied as an $n$ vector of ${1..n}$  then $\mathcal P_t = f_{\mathcal M}(V_t)$ where $\mathcal P_t$ is the $t^{th}$ element of the permutation, $f_{\mathcal M}$ is an arbitrary function that specifies which node to choose next based off of the current graph and the model parameters $\mathcal M$, and $V_t$ is the state of the agents at time $t$. The optimal schedule is the schedule that maximizes $J = \sum_{i=1}^n \mathbb E[\mathbb I(v_{in} = Y)]$ with respect to the schedule $\mathcal P$, where $\mathbb I$ is the indicator function and $v_{it}$ is the state of agent $i$ at time $t$. Before the game is played each agent also gets a prefered type which is drawn from a bionmial where $p(type_i = Y) = p < \frac{1}{2}$. The utility of a node $i$ at time $t$ is $u_{it}(v_{it}) = \pi \mathbb I(v_{it} = type_i) + \sum_{j \in \mathcal N(i)} \mathbb I(v_{it} = v_{jt}), v_{it} \in \{Y, N\}$, where $\pi$ is a constant that determines how much utility an agent gets from choosing its prefered type. If an agent is deciding its type at time $t$ a myopic agent $i$ will choose to maxmimize its current utility $\arg\max_{v_{it} \in \{Y, N\}} u_{it}$ where as a strategic agent will choose to maximize its expected final utility $\arg\max_{v_{it} \in \{Y, N\}} \mathbb E[u_{in}]$. We are interested in looking at the difference between the optimal number of minority choosing agents when the agents are strategic versus myopic ($J_{myopic} / J_{strategic}$).

Ongoing work by Martin et al. suggests that the scheduler payoff is always lower in the strategic case than in the myopic case, but we are interested quantifying this difference. To tackle this problem we propose investigating efficient algorithms for finding the optimal scheduling strategies on specific classes of graphs, which we hope will provide insight into quantifying the difference between the expected number of agents picking the minority type when the agents are strategic versus myopic.

We believe that an efficient dynamic programming algorithm exists for finding the optimal schedule on graphs that can be specified as a $2$-block graph. We define a $k$-block graph as any graph whose adjacency matrix can be represented as symmetric block matrix with $k^2$ blocks such that each block is either all $1$'s or all $0$'s. The idea behind the algorithm is to do a pruned backward induction from potential termination states. Because agents in a single block are substitutable, the potential choice at each round is significantly reduced. This combined with pruning should provide the necessary requirements to make this algorithm efficient. Work by Chierichetti et al.\cite{Chierichetti12} has already bounded the optimal schedule payout for arbitrary graphs, however applying an algorithm similar to ours for strategic agents should be no worse in complexity for myopic agents. In addition to the optimal offline schedule, the inductive nature of this algorithm should mean it also calculates the optimal online schedule. This efficient optimal algorithm should allow us to calculate the difference between strategic and myopic agents under offline and online schedules, and will hopefully give insight into bounds on the difference for arbitrary 2-block graphs.

There are a number of possible extensions for this algorithm that should also broaden the class of graphs that our results could apply to. These extensions include scaling to $k$-block graphs, directed graphs, and stochastic block graphs. Extending the algorithm to $k$-block graphs shouldn’t conceptually change the algorithm, but we believe that the original algorithm may scale exponentially in $k$ for a $k$-block graph. However, for bounded $k$, the complexity is still be moderately efficient. Additionally, scaling the algorithm up to $k$-blocks may expose room for improvement in the algorithm and provide further insight into quantifying the difference between myopic and strategic agents in this setting. Modifying the algorithm to account for directed graphs as long as they still fit into a blockmodel is also likely trivial and would further increase the number of graphs that we can analyze theoretically although the applicability of directed graphs is not immediately apparent. Another potentially easy extension of this would be to look at stochastic blockmodels\cite{Wang87, Snijders97}. It seems reasonable that the algorithm for static block graph models might extend in expectation to unrealized (and hence offline scheduled) stochastic blockmodels. This extension could likely never extend to realized graphs or to online schedules, but stochastic blockmodels are much closer to graphs found experimentally than $k$-block graphs.

\bibliographystyle{plain}
\bibliography{references}
\end{document}